---
title: "bechdel-proj"
output: html_document
date: "2024-03-07"
authors: "Alexa Kaplan, Alec Beland, Catherine Crowell, Carolyn Kelly"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      message = F,
                      warning = F)

#load packages we will need
install.packages("dplyr")
pacman::p_load(tidyverse)

#we can also change default themes here if necessary

#read in the data file here!!!

bechdel <- read_csv("bechdel2.csv")

```

I. Introduction: At the beginning of your paper, you must describe the data, in a paragraph. Note the following:


The Bechdel Test assesses female representation in film. To pass, the film must show two women who talk to each other about something other than men. The data we are working with is the "Movies Bechdel Test" data set from the miscellaneous data sets provided on Brightspace. It contains data on 1,793 films ranging from the years 1970 to 2013, including their Bechdel scores on a scale of 1 to 5. A score of 1 passes the test; a score of 2 is "dubious" as it is unclear whether the dialogue between the two women is sufficient to pass; a film with a 3 includes women who talk to each other exclusively about men; a 4 has women who do not talk to each other; and, finally, a 5 has no female characters. 

Sampling bias is unlikely since the films since the sample is large at almost 1,800 films. There should be no bias in measurements here since the profits are quantitative, not qualitative, measurements, as well as the Bechdel Test which turns objective criteria into a numeric score.

This data is of interest because, while we like to think gender equality and gender roles have drastically improved in the past five decades (range of this data set), we can look at the variable of female representation in film to check social progress. 

As for data cleaning, data points with missing values were removed so that results would not be skewed. 

 What is the source of the data? Where and when was it created?
 If it is a sample, from what population was it drawn, and how was the sample selected?
 Do you suspect any sampling bias?
 Was it an experiment or an observational study?
 How were measurements taken, or questions asked?
 Do you suspect any bias in the questions or measurements?
 Why is this data of interest to you, and why should the class find it interesting?
 What kind of data cleaning was necessary (R code for this must show…)
(^^Just for reference, will be deleted^^)

## Introduction

The **Bechdel Test** ...


```{r question 1 - Introduction}
#data cleaning here described above...
bechdel_factored <- bechdel |> 
  mutate(score = case_when(Btest5 == "1 ok" ~ 1,
                           Btest5 == "2 dubious" ~ 2,
                           Btest5 == "3 men" ~ 3,
                           Btest5 == "4 notalk" ~ 4,
                           Btest5 == "5 nowomen" ~ 5)) |> 
  select(-Btest5)
  
```

II. Data Visualizations: Write R code to create some relevant descriptive graphs, using techniques that
we’ve used in class (ggplot, maybe dplyr). About 4 or 5 graphs should be plenty, depending on
complexity.
For each graph and numerical summary, write a paragraph or two summarizing what you see, and
suggesting some implications. For example, describe patterns that you observe in a graph, and
suggest why they make sense, given what you know about the subject, or if they are unexpected.
Do you think there is a cause-effect relationship between any variables? Explain your reasoning.  
(^^Just for reference, will be deleted^^)

## Data Visualizations


```{r question 2 - Data Visualizations}
# line to rating by score 
bechdByRating <-
  ggplot(
    data = bechdel_factored,
    mapping = aes(
      x = averating,
      y = score
    )
  ) +
  geom_line()

bechdHist <-
  ggplot(
    data = bechdel_factored,
    mapping = aes(x = score)
  ) +
  geom_histogram() +
  scale_x_reverse()
# scatter to show rating by score with lines
bechdScatter <- ggplot(
  data = bechdel_factored,
  mapping = aes(
    x = averating,
    y = budget13,
    color = factor(score)
  )
) +
  geom_point() +
  
  geom_smooth(
    method = loess,
    formula = y~x,
    se = F
  )

# side-by-side bars to show domestic vs international profit

# compare year to bechdel rating


# show graphs
bechdByRating
bechdHist
bechdScatter
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
