---
title: "bechdel-proj"
output: html_document
date: "2024-03-07"
authors: "Alexa Kaplan, Alec Beland, Catherine Crowell, Carolyn Kelly"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      message = F,
                      warning = F)

#load packages we will need
install.packages("dplyr")
pacman::p_load(tidyverse)

#we can also change default themes here if necessary

#read in the data file here

bechdel <- read_csv("bechdel2.csv")

```

I. Introduction: At the beginning of your paper, you must describe the data, in a paragraph. Note the following:
 What is the source of the data? Where and when was it created?
 If it is a sample, from what population was it drawn, and how was the sample selected?
 Do you suspect any sampling bias?
 Was it an experiment or an observational study?
 How were measurements taken, or questions asked?
 Do you suspect any bias in the questions or measurements?
 Why is this data of interest to you, and why should the class find it interesting?
 What kind of data cleaning was necessary (R code for this must show…)
(^^Just for reference, will be deleted^^)

## Introduction

The **Bechdel Test** ...


```{r question 1 - Introduction}
#data cleaning here described above...
```

II. Data Visualizations: Write R code to create some relevant descriptive graphs, using techniques that
we’ve used in class (ggplot, maybe dplyr). About 4 or 5 graphs should be plenty, depending on
complexity.
For each graph and numerical summary, write a paragraph or two summarizing what you see, and
suggesting some implications. For example, describe patterns that you observe in a graph, and
suggest why they make sense, given what you know about the subject, or if they are unexpected.
Do you think there is a cause-effect relationship between any variables? Explain your reasoning.  
(^^Just for reference, will be deleted^^)

## Data Visualizations


```{r question 2 - Data Visualizations}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
